Towards Practical First-Order Model Counting

### 2025-01-24 doc/paper/paper.tex
Resize the figure

% 1. Introduction
% Objective: Briefly state the goals of the experiments.
% Overview: Summarize the structure of the section.

% 2. Experimental Setup

% Baselines: List and describe the baseline methods or systems used for
% comparison.

@@

% It would not make sense to include the original \textsc{Crane} algorithm in
% the experiments or evaluate \Cranetwo{} without the C++ code generation
% described in \cref{sec:cpp}. Indeed, these algorithms can only produce
% mathematical definitions of functions without evaluating them.

% While greedy search is more efficient (in terms of being able to handle larger
% formulas), it is a heuristic that can miss some solutions.

% Datasets: Provide details about the datasets used in the experiments,
% including size, source, and any preprocessing steps.

@@

%Since \textsc{FastWFOMC} does not support many-sorted logic, our experiments are
%limited to formulas with a single domain. However, many of the counting problems
%used in the experimental evaluation of
%\textsc{Crane}~\cite{DBLP:conf/kr/DilkasB23} become equivalent (in terms of
%generating equivalent integer sequences) when restricted to a single domain.
%Additionally, comparing \Cranetwo{} and \textsc{FastWFOMC} on a broader set of
%problems is challenging because each problem needs to be represented in two
%(quite different) logics: \FO{} and \UFO{}.

@@

% Metrics: Define the performance metrics used to evaluate the results (e.g.,
% accuracy, precision, recall, F1-score, runtime).

%A considerable difference between \textsc{ForcLift} and the other two algorithms
%is that \textsc{ForcLift} does not support arbitrary precision whereas the other
%algorithms use the GNU Multiple Precision Arithmetic Library. Thus, when the
%count exceeds finite precision, \textsc{ForcLift} returns $\infty$.

@@

% Environment: Describe the hardware and software environment (e.g.,
% specifications of the computer, operating system, libraries, and tools used).

% 3. Results and Analysis
% Presentation of Results: Use tables, graphs, and charts to present the experimental results clearly.
% Comparison with Baselines: Compare your method’s results with the baselines.
% Discussion: Analyze and interpret the results, highlighting the strengths and potential weaknesses of your approach.
% Insights: Provide insights or patterns observed from the results.
% Challenges: Discuss any challenges or unexpected outcomes.
% Acknowledgement: Honestly acknowledge any limitations of your experiments or method.

@@

\todo[inline,caption={}]{
  \begin{itemize}
    \item On the `friends' and `functions' benchmarks, \textsc{ForcLift} runs
          until the model count exceeds $2^{31}-1$.
    \item We are not aware of any formulas on which \Cranetwo{} scales worse
          compared to either \textsc{ForcLift} or \textsc{FastWFOMC}. The one
          advantage that \textsc{FastWFOMC} has over \Cranetwo{} is its support
          for counting quantifiers.
    \item Regarding programming languages and accuracy, we verified that the
          answers match for smaller domain sizes. Also, although written in
          different languages, both \Cranetwo{} and \textsc{FastWFOMC} use the
          GNU Multiple Precision Arithmetic Library.
  \end{itemize}
}

\Cref{fig:plot} presents a summary of the experimental results. Only
\textsc{FastWFOMC} and \Cranebfs{} could handle the bijection-counting problem.
For this benchmark, the largest domain sizes these algorithms could accommodate
were \num{64} and \num{4096}, respectively. On the other two benchmarks,
\textsc{ForcLift} had the lowest runtime. However, due to its finite precision,
it only scaled up to domain sizes of \num{16} and \num{128} for `friends' and
`functions', respectively. \textsc{FastWFOMC} outperformed \textsc{ForcLift} in
the case of `friends', but not `functions', as it could handle domains of size
\num{1024} and \num{64}, respectively. Furthermore, both \Cranebfs{} and
\Cranegreedy{} performed similarly on both benchmarks. Similarly to the
`bijections' benchmark, \Cranetwo{} significantly outperformed the other two
algorithms, scaling up to domains of size \num{8192} and \num{67108864},
respectively.

@@

% 8. Conclusion
% Summary: Briefly summarize the key findings from the experimental results.
% Future Work: Suggest directions for future research based on your findings.

@@

In this work, we have presented a scalable automated FOKC-based approach to
FOMC. Our algorithm involves completing the definitions of recursive functions
and subsequently translating all function definitions into C++ code. Empirical
results demonstrate that \Cranetwo{} can scale to larger domain sizes than
\textsc{FastWFOMC} while supporting a wider range of formulas than
\textsc{ForcLift}. The ability to efficiently handle large domain sizes is
particularly crucial in the weighted setting, as illustrated by the `friends'
example discussed in \cref{sec:experiments}, where the model captures complex
social networks with probabilistic relationships. Without this scalability, the
practical usefulness of these models would be limited.

>>>>

@@

@@

@@

@@

@@

\todo[inline,caption={}]{
  \begin{itemize}
     \item We are not aware of any formulas on which \Cranetwo{} scales worse
          compared to either \textsc{ForcLift} or \textsc{FastWFOMC}. The one
          advantage that \textsc{FastWFOMC} has over \Cranetwo{} is its support
          for counting quantifiers.
    \item Regarding programming languages and accuracy, we verified that the
          answers match for smaller domain sizes. Also, although written in
          different languages, both \Cranetwo{} and \textsc{FastWFOMC} use the
          GNU Multiple Precision Arithmetic Library.
  \end{itemize}
}

\Cref{fig:plot} presents a summary of the experimental results. Only
\textsc{FastWFOMC} and \Cranebfs{} could handle the bijection-counting problem.
For this benchmark, the largest domain sizes these algorithms could accommodate
were \num{64} and \num{4096}, respectively. On the other two benchmarks,
\textsc{ForcLift} had the lowest runtime. However, since it can only handle
model counts smaller than $2^{31}$, it only scales up to domain sizes of
\num{16} and \num{128} for `friends' and `functions', respectively.
\textsc{FastWFOMC} outperformed \textsc{ForcLift} in the case of `friends', but
not `functions', as it could handle domains of size \num{1024} and \num{64},
respectively. Furthermore, both \Cranebfs{} and \Cranegreedy{} performed
similarly on both benchmarks. Similarly to the `bijections' benchmark,
\Cranetwo{} significantly outperformed the other two algorithms, scaling up to
domains of size \num{8192} and \num{67108864}, respectively.

@@

@@

In this work, we have presented a scalable automated FOKC-based approach to
FOMC\@. Our algorithm involves completing the definitions of recursive functions
and subsequently translating all function definitions into C++ code. Empirical
results demonstrate that \Cranetwo{} can scale to larger domain sizes than
\textsc{FastWFOMC} while supporting a wider range of formulas than
\textsc{ForcLift}. The ability to efficiently handle large domain sizes is
particularly crucial in the weighted setting, as illustrated by the `friends'
example discussed in \cref{sec:experiments}, where the model captures complex
social networks with probabilistic relationships. Without this scalability, the
practical usefulness of these models would be limited.
